---
title: "Milestone 4 Regression Tree"
author: "Group 13"
date: "`r Sys.Date()`"
output: pdf_document
---

Regression Question: How indicative are the prices of everyday commodities of a city’s average income?

```{r}
library(pacman)
p_load(MASS, tree, randomForest, gbm, dplyr, tidyverse)

train <- readRDS("C:/Users/abz20/OneDrive/Desktop/UVA Courses/Statistical Machine Learning (STAT 4630)/STAT4630Project/Scripts/train.rds")
test <- readRDS("C:/Users/abz20/OneDrive/Desktop/UVA Courses/Statistical Machine Learning (STAT 4630)/STAT4630Project/Scripts/test.rds")
```

Parts A and B: Data Cleaning/Processing, Subsetting to Plausible Predictors
```{r}
# exclude predictors related to each other or unrelated to question
train = train %>% dplyr::select(-city,
                                -country,
                                -beer.rest.domestic,
                                -beer.rest.imported,
                                -coffee,
                                -soda,
                                -water.rest,
                                -taxi.km,
                                -taxi.hr,
                                -rent1.center,
                                -rent1.outer,
                                -rent3.center,
                                -rent3.outer,
                                -sqm.center,
                                -sqm.outer,
                                -quality)

test = test %>% dplyr::select(-city,
                                -country,
                                -beer.rest.domestic,
                                -beer.rest.imported,
                                -coffee,
                                -soda,
                                -water.rest,
                                -taxi.km,
                                -taxi.hr,
                                -rent1.center,
                                -rent1.outer,
                                -rent3.center,
                                -rent3.outer,
                                -sqm.center,
                                -sqm.outer,
                                -quality)
```

Excluded variables:
- city and country because they're identifiers, neither predictor nor response
- beer.rest.domestic, beer.rest.imported, coffee, soda, water.rest, because they're related to meal1, meal2, and mcmeal
- taxi.km and taxi.hr because they're related to taxi.start
- rent1.center because this was converted to a categorical variable, expensive, so it would be redundant to include this
- rent1.outer, rent3.center, rent3.outer, sqm.center, and sqm.outer because they're closely related to "expensive"
- quality because this isn't related to cost of living

Part C: Recursive Binary Splitting
```{r}
# fit tree model using training data with binary recursive splitting
tree.result = tree::tree(salary ~ ., data = train) # x54 is monthly salary
```

i. 
```{r}
# see output
summary(tree.result)
```

ii. 8 terminal nodes

iii. Variables actually used in tree construction: "cinema" "ticket" "meal1" "bread" "cigarettes" "pass" "preschool" 

iv. 
```{r}
# decision tree built on training data with recursive binary splitting
plot(tree.result)
text(tree.result, cex=0.4)
```

v. This answers our question of interest by selecting which predictors out of all our plausible predictors are most important in predicting the average monthly salary in a city. 

vi. 
```{r}
# find predictions for test data
tree.pred.test = predict(tree.result, newdata=test) 

# find test MSE
recursive_binary_test_mse = mean((test$salary - tree.pred.test)^2)

cat("Recursive Binary Test MSE:", recursive_binary_test_mse)
```

Part D: Pruned Tree

Note: The pruned tree is the same as the tree from recursive binary
splitting. I still answered the questions from Part C just in case we decide to redo this. 

```{r}
# use 10-fold CV to prune tree
cv.Dataset = tree::cv.tree(tree.result, K=10)
cv.Dataset

# plot of residual mean deviance vs size of tree with pruning
plot(cv.Dataset$size, cv.Dataset$dev, type="b", xlab="Size of Tree", ylab="Res Mean Deviance")

# see size of tree which gives best tree based on pruning and 10-fold CV
trees.num = cv.Dataset$size[which.min(cv.Dataset$dev)]
trees.num
```

```{r}
# refit with training data
tree.train = tree::tree(salary ~ ., data = train)
prune.train = tree::prune.tree(tree.train, best=trees.num)

# decision tree with pruning, with training data
plot(prune.train)
text(prune.train, cex=0.4)

# numerical summary of pruned tree
prune.train
```


i. 
```{r}
# see output
summary(prune.train)
```

ii. 8 terminal nodes

iii. Variables actually used in tree construction: "cinema" "ticket" "meal1" "bread" "cigarettes" "pass" "preschool"

iv. 
```{r}
# decision tree built on training data with recursive binary splitting
plot(prune.train)
text(prune.train, cex=0.4)
```

v. This answers our question of interest by selecting which predictors out of all our plausible predictors are most important in predicting the average monthly salary in a city, but with less overfitting.  

vi. 
```{r}
# find predictions for test data
tree.pred.test = predict(prune.train, newdata=test) 

# find test MSE
pruned_tree_test_mse = mean((test$salary - tree.pred.test)^2)

cat("Pruned Tree Test MSE:", pruned_tree_test_mse)
```

Part E: Random Forests
```{r}
rf.class = randomForest::randomForest(salary ~ ., data=train, mtry=2,importance=TRUE) # mtry = p/3 for regression tress (see Lab 8b)
rf.class

importance(rf.class)
varImpPlot(rf.class)
```

i. ticket and cinema are most important predictors

```{r}
# test accuracy with Random Forest
pred.rf<-predict(rf.class, newdata=test)

RF_test_mse = mean((test$salary - pred.rf)^2)

cat("Random Forest Test MSE:", RF_test_mse)
```
Part F: Conclusion 

i. 
```{r}
cat("Recursive Binary Test MSE:", recursive_binary_test_mse, "\n",
    "Pruned Test MSE:", pruned_tree_test_mse, "\n",
    "Random Forest Test MSE:", RF_test_mse)
```




























